### 2020-08-06 paper/app-addmaterial.tex
Fix low-invariant mass slices of CMS DY

\section{Additional material}
\label{app:add_plots}

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/pineappl_ATLASZHIGHMASS49FB_lowstat}\\
    \caption{Low-statistics ATLAS high-mass Drell--Yan.}
    \label{fig:atlaszhighmass49fb-lowstat}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/pineappl_CMSDY2D11_bin1}%
    \includegraphics[width=0.5\textwidth]{figures/pineappl_CMSDY2D11_bin2}
    \caption{PineAPPL comparison for CMS 2D Drell--Yan.}
    \label{fig:cmsdy2d11_bins12}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/pineappl_ATLAS_TTB_DIFF_8TEV_LJ_TRAP}%
    \caption{PineAPPL comparison for top pair.}
    \label{fig:cmsdy2d11_bins56}
\end{figure}