### 2020-08-06 paper/app-lumis.tex
First try to appendix B

\section{Parton Luminosities}
\label{app:lumis}

\begin{table}[!t]
  \centering
  \tiny
  \input{tables/TableB1.tex}\\
  \caption{The 34 parton luminosities contributing to the predictions, accurate
    to NLO QCD+EW, of the Drell--Yan lepton-pair production measured by the
    ATLAS experiment at a centre-of-mass energy of
    \SI{7}{\tera\electronvolt}~\cite{Aad:2013iua}. The ten largest contributions
    to the cross section (irrespective of their sign) are reported, in
    percentage, for each bin. Empty spaces typically denote contributions
    smaller than a fraction of percent. The bin edges are in the invariant mass
    of the lepton pair, $M_{\ell\bar\ell}$, as follows: 116, 130, 150, 170, 190,
    210, 230, 250, 300, 400, 500, 700, 1000, 1500 [GeV].
  }
  \label{tab:B1}
\end{table}