### 2020-08-13 paper/paper.tex
Acknowledgements

\documentclass[a4paper,11pt]{article}
\pdfoutput=1
\pdfsuppresswarningpagegroup=1

\usepackage[UKenglish]{babel}
\usepackage{booktabs}
\usepackage{csquotes}
\usepackage{jheppub}
\usepackage[T1]{fontenc}
\usepackage{listings}
\usepackage{lmodern}
\usepackage[binary-units=true,detect-weight=true,table-format=4.2]{siunitx}
\usepackage{xcolor}
\usepackage[percent]{overpic}
\usepackage{mathcomp}
\usepackage{tabularx}
\usepackage{textcomp}
\usepackage{xltabular}

\newcommand{\alphas}{\alpha_\mathrm{s}}
\DeclareSIUnit\permille{\text{\textperthousand}}

\preprint{Nikhef-2020-026 TIF-UNIMI-2020-22}

\title{PineAPPL:\\ combining EW and QCD corrections for PDF fits}

\author[a]{S. Carrazza,}
\author[b]{E.R. Nocera,}
\author[a]{C. Schwan}
\author[a]{and M. Zaro}

\affiliation[a]{Tif Lab, Dipartimento di Fisica, 
Universit\`a di Milano and INFN, Sezione di Milano, 20133 Milano, Italy}
\affiliation[b]{Nikhef Theory Group, Science Park 105, 1098 XG Amsterdam, 
The Netherlands}

\emailAdd{stefano.carrazza@mi.infn.it}
\emailAdd{e.nocera@nikhef.nl}
\emailAdd{christopher.schwan@mi.infn.it}
\emailAdd{marco.zaro@mi.infn.it}

\abstract{We introduce \textsc{PineAPPL}, a computer library allowing to produce fast-in\-ter\-po\-la\-tion
grids for fitting parton distribution functions (PDFs)
including corrections of strong and electroweak origin. After presenting
the features and performance of such a code, we show how to interface \textsc{PineAPPL} with general-purpose
softwares that allow the user to compute such corrections, providing an explicit example with
\textsc{MadGraph5\_aMC@NLO}. We showcase some phenomenological result focusing on processes and observables
relevant for PDF fits, and 
we discuss what steps need to be undertaken in the presentation of experimental data in order to have them
consistently employed in a fit including electroweak corrections.}

\begin{document}

\maketitle
\flushbottom

\input{sec-introduction}
\input{sec-pineappl}
\input{sec-results}
\input{sec-doublecounting}
\input{sec-conclusion}

\appendix

\acknowledgments
We are grateful to Rikkert Frederix for discussions about
\textsc{aMCfast}, and to Stefan Dittmaier for a critical reading of the
manuscript. We acknowledge discussions with 
Florencia Canelli, Stefano Camarda, Paolo Francavilla, Abideh Jafari, Andreas Jung, Elizaveta Shabalina, Wolfgang Wagner about the 
treatment of EW effects in experimental data.
%Alessandro Vicini\\
S.C.\ and C.S.\ are supported by the European Research Council under the European Union's
Horizon 2020 research and innovation Programme (grant agreement no.\ 740006).
E.R.N.\ is supported by the European Commission through a Marie
Sk\l odowska-Curie Action (grant agreement no.\ 752748).
M.Z.\ thanks the Nikhef institute in Amsterdam,
where he was employed when this paper was started.

\input{app-pineappl}
\input{app-lumis}
\input{app-addmaterial}

\bibliographystyle{JHEP}
\bibliography{paper}

\end{document}