### 2020-08-13 paper/paper.tex
Abstract

\documentclass[a4paper,11pt]{article}
\pdfoutput=1
\pdfsuppresswarningpagegroup=1

\usepackage[UKenglish]{babel}
\usepackage{booktabs}
\usepackage{csquotes}
\usepackage{jheppub}
\usepackage[T1]{fontenc}
\usepackage{listings}
\usepackage{lmodern}
\usepackage[binary-units=true,detect-weight=true,table-format=4.2]{siunitx}
\usepackage{xcolor}
\usepackage[percent]{overpic}
\usepackage{mathcomp}
\usepackage{tabularx}
\usepackage{textcomp}
\usepackage{xltabular}

\newcommand{\alphas}{\alpha_\mathrm{s}}
\DeclareSIUnit\permille{\text{\textperthousand}}

\preprint{Nikhef-2020-026 TIF-UNIMI-2020-22}

\title{PineAPPL:\\ combining EW and QCD corrections for PDF fits}

\author[a]{S. Carrazza,}
\author[b]{E.R. Nocera,}
\author[a]{C. Schwan}
\author[a]{and M. Zaro}

\affiliation[a]{Tif Lab, Dipartimento di Fisica, 
Universit\`a di Milano and INFN, Sezione di Milano, 20133 Milano, Italy}
\affiliation[b]{Nikhef Theory Group, Science Park 105, 1098 XG Amsterdam, 
The Netherlands}

\emailAdd{stefano.carrazza@mi.infn.it}
\emailAdd{e.nocera@nikhef.nl}
\emailAdd{christopher.schwan@mi.infn.it}
\emailAdd{marco.zaro@mi.infn.it}

\abstract{We introduce \textsc{PineAPPL}, a library that produces
  fast-interpolation grids of physical cross sections, computed with a 
  general-purpose Monte Carlo generator, accurate to fixed order in the
  strong, electroweak, and combined strong-electroweak couplings. We
  discuss its features in comparison to similar software
  available in the literature, and we explicitly interface it to
  \textsc{MadGraph5\_aMC@NLO}. We validate the performance of \textsc{PineAPPL}
  by computing fast-interpolation grids, accurate to next-to-leading order in
  the strong and electroweak couplings, for a representative set of processes
  commonly used in the determination of the parton distribution functions (PDFs)
  of the proton. We formulate a recommendation on the format of the experimental
  deliverables in order to consistently incorporate them in forthcoming
  fits of PDFs including electroweak corrections.}

\begin{document}

\maketitle
\flushbottom

\input{sec-introduction}
\input{sec-pineappl}
\input{sec-results}
\input{sec-doublecounting}
\input{sec-conclusion}

\appendix

\acknowledgments
We are grateful to Rikkert Frederix for discussions about
\textsc{aMCfast}, and to Stefan Dittmaier for a critical reading of the
manuscript. We acknowledge discussions with 
Florencia Canelli, Stefano Camarda, Paolo Francavilla, Abideh Jafari, Andreas Jung, Elizaveta Shabalina, Wolfgang Wagner about the 
treatment of EW effects in experimental data.
%Alessandro Vicini\\
S.C.\ and C.S.\ are supported by the European Research Council under the European Union's
Horizon 2020 research and innovation Programme (grant agreement no.\ 740006).
E.R.N.\ is supported by the European Commission through a Marie
Sk\l odowska-Curie Action (grant agreement no.\ 752748).
M.Z.\ thanks the Nikhef institute in Amsterdam,
where he was employed when this paper was started.

\input{app-pineappl}
\input{app-lumis}
\input{app-addmaterial}

\bibliographystyle{JHEP}
\bibliography{paper}

\end{document}