### 2020-08-25 paper/paper.tex
Mention PDFs in the abstract; emphasise multi-coupling expansion in abstract and sec.1; discuss choice of processes in sec.3

\documentclass[a4paper,11pt]{article}
\pdfoutput=1
\pdfsuppresswarningpagegroup=1

\usepackage[UKenglish]{babel}
\usepackage{booktabs}
\usepackage{csquotes}
\usepackage{jheppub}
\usepackage[T1]{fontenc}
\usepackage{listings}
\usepackage{lmodern}
\usepackage[binary-units=true,detect-weight=true,digitsep=none,table-format=4.2]{siunitx}
\usepackage{xcolor}
\usepackage[percent]{overpic}
\usepackage{mathcomp}
\usepackage{tabularx}
\usepackage{textcomp}
\usepackage{xltabular}

\newcommand{\alphas}{\alpha_\mathrm{s}}
\DeclareSIUnit\permille{\text{\textperthousand}}

\preprint{Nikhef-2020-026 TIF-UNIMI-2020-22}

\title{PineAPPL: combining EW and QCD corrections\\
  for fast evaluation of LHC processes}

\author[a]{S. Carrazza,}
\author[b]{E.R. Nocera,}
\author[a]{C. Schwan}
\author[a]{and M. Zaro}

\affiliation[a]{Tif Lab, Dipartimento di Fisica, 
Universit\`a di Milano and INFN, Sezione di Milano, 20133 Milano, Italy}
\affiliation[b]{Nikhef Theory Group, Science Park 105, 1098 XG Amsterdam, 
The Netherlands}

\emailAdd{stefano.carrazza@mi.infn.it}
\emailAdd{e.nocera@nikhef.nl}
\emailAdd{christopher.schwan@mi.infn.it}
\emailAdd{marco.zaro@mi.infn.it}

\abstract{We introduce \textsc{PineAPPL}, a library that produces
  fast-interpolation grids of physical cross sections, computed with a 
  general-purpose Monte Carlo generator, accurate to fixed order in the
  strong, electroweak, and combined strong-electroweak couplings. We
  demonstrate this unique ability, that distinguishes \textsc{PineAPPL} from
  similar software available in the literature, by interfacing it to
  \textsc{MadGraph5\_aMC@NLO}. We compute fast-interpolation grids,
  accurate to next-to-leading order in the strong and electroweak couplings,
  for a representative set of LHC processes for which EW corrections may have
  a sizeable effect on the accuracy of the corresponding theoretical
  predictions. We formulate a recommendation on the format of the experimental
  deliverables in order to consistently compare them with computations
  that incorporate EW corrections, and specifically to determine
  parton distribution functions to the same accuracy.}

\begin{document}

\maketitle
\flushbottom

\input{sec-introduction}
\input{sec-pineappl}
\input{sec-results}
\input{sec-doublecounting}
\input{sec-conclusion}

\appendix

\acknowledgments
We would like to thank Stefan Ditt\-mai\-er, Stefano Forte, Zahari Kassabov, and Juan Rojo for a critical reading of the manuscript, and we are grateful to Rikkert Frederix for discussions about \textsc{aMCfast}.
We acknowledge discussions with
Florencia Canelli, Stefano Camarda, Paolo Francavilla, Abideh Jafari, Andreas Jung, Elizaveta Shabalina, Wolfgang Wagner about the 
treatment of EW effects in experimental data.
S.C.\ and C.S.\ are supported by the European Research Council under the European Union's
Horizon 2020 research and innovation Programme (grant agreement no.\ 740006).
E.R.N.\ is supported by the European Commission through a Marie
Sk\l odowska-Curie Action (grant agreement no.\ 752748).
M.Z.\ thanks the Nikhef institute in Amsterdam,
where he was employed when this paper was started.

\input{app-pineappl}
\input{app-lumis}
\input{app-addmaterial}

\bibliographystyle{JHEP}
\bibliography{paper}

\end{document}