### 2020-10-14 paper/answer_to_the_editor.txt
added comment on point 10

Dear Editor,

We would like to thank the referee for carefully reading the manuscript and
making useful comments and suggestions.

In the following our answers to the issues raised by the referee:

- Issues 1, 3, 4, 5, 6, 8, 9, and 11 have all been addressed as suggested or
  corrected otherwise.
- Issue 2: The decomposition of the fixed-order expansion written down in eq.
  (2.1) holds to all orders in both the strong and electroweak coupling, and
  therefore the mentioned sentence is correct—it only concerns PineAPPL. The
  Monte Carlo that we used is tied to NLO, of course.
- Issue 7:
- Issue 10: A brief discussion on this matter has been added at the beginning
  of section 3. The reference suggested (Buccioni:2020cfi), is cited in section 
  3.2.1, in the discussion relevant for Drell-Yan.
- Issue 12: Although we agree with the referee that the mentioned tables are
  interesting only to a few readers, the main text refers to it in a few
  places and therefore removing it from the main document would require bigger
  changes. We see no harm in leaving the tables in the document and therefore
  decided to keep them.

Note that all changes have been highlighted in red in the text.

With these changes, we hope that the current version will meet the criteria of
publication of JHEP.

Best regards,
The authors