### 2020-10-26 paper/answer_to_the_editor.txt
Update answer to the editor

Dear Editor,

We would like to thank the referee for carefully reading the manuscript and
making useful comments and suggestions.

For convenience all changes are marked in the pre-compiled document as
red/strikethrough (removed text) and blue/underlined (added text).

In the following our answers to the issues raised by the referee:

- Issues 1, 3, 4, 5, 6, 8, 9, and 11 have all been addressed as suggested, or
  corrected otherwise.
- Issue 2: The decomposition of the fixed-order expansion written down in eq.
  (2.1) holds to all orders in both the strong and electroweak coupling, and
  therefore the mentioned sentence is correct, as it only concerns PineAPPL.
  The Monte Carlo that we used is limited to NLO, of course.
- Issue 7: For Drell-Yan we checked that a dynamical scale set to Mll does
  change the NLO QCD in the largest invariant mass bins and therefore also the
  scale uncertainty. In comparison to the shown data uncertainty the changes
  are very minor, and therefore we did not update plots and text.
  For the Z transverse momentum process we did update the both plots and text,
  where we indeed see a significant increase of the scale variation band in
  the large transverse momentum bins. The electroweak corrections are also
  smaller, since we normalise to NLO QCD.
- Issue 10: A brief discussion on this matter has been added at the beginning
  of section 3. The reference suggested (Buccioni:2020cfi), is cited in section 
  3.2.1, in the discussion relevant for Drell-Yan.
- Issue 12: Although we agree with the referee that the mentioned tables are
  interesting only to a few readers, the main text refers to it in a few
  places and therefore removing it from the main document would require bigger
  changes. We see no harm in leaving the tables in the document and therefore
  decided to keep them.

With these changes, we hope that the current version will meet the criteria of
publication of JHEP.

Best regards,
The authors