### 2021-03-30 paper2/paper.tex
Change title

\documentclass[a4paper,11pt]{article}
\pdfoutput=1
\pdfsuppresswarningpagegroup=1

% load before csquotes
\usepackage[utf8]{inputenc}

\usepackage[UKenglish]{babel}
%\usepackage{booktabs}
%\usepackage{csquotes}
\usepackage{jheppub}
\usepackage[T1]{fontenc}
%\usepackage{listings}
\usepackage{lmodern}
\usepackage[binary-units=true,detect-weight=true,group-digits=false,table-format=4.2]{siunitx}
%\usepackage{xcolor}
%\usepackage[percent]{overpic}
%\usepackage{mathcomp}
%\usepackage{tabularx}
%\usepackage{textcomp}

%\newcommand{\alphas}{\alpha_\mathrm{s}}
%\DeclareSIUnit\permille{\text{\textperthousand}}

\preprint{TODO}

\title{TODO}

\author[a]{S. Carrazza,}
\author[b]{E.R. Nocera,}
\author[a]{C. Schwan}
\author[a]{and M. Zaro}

\affiliation[a]{Tif Lab, Dipartimento di Fisica, 
Universit\`a di Milano and INFN, Sezione di Milano, 20133 Milano, Italy}
\affiliation[b]{Nikhef Theory Group, Science Park 105, 1098 XG Amsterdam, 
The Netherlands}

\emailAdd{stefano.carrazza@mi.infn.it}
\emailAdd{e.nocera@nikhef.nl}
\emailAdd{christopher.schwan@mi.infn.it}
\emailAdd{marco.zaro@mi.infn.it}

\abstract{TODO}

\arxivnumber{TODO}

\begin{document}

\maketitle
\flushbottom

\input{sec-introduction}
\input{sec-results}
\input{sec-datasets}
\input{sec-conclusion}

\appendix

\acknowledgments
S.C.\ and C.S.\ are supported by the European Research Council under the European Union's Horizon 2020 research and innovation Programme (grant agreement no.\ 740006).

\input{app-misc}

\bibliographystyle{JHEP}
\bibliography{paper}

\end{document}